\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

This thesis would not have been as easy and enjoyable to put together without the help of the people mentioned below.
%Natalie Cooper for kind of everything
It would have not been possible at all, however, without the devotion of my supervisor Natalie Cooper.
Thank you for the countless time you spend patiently helping/coaching/teaching/mentoring me but also for being always motivated and positive about this whole project (especially when it was going as planned).
I've learned a lot from working with you both in terms of skills and of the attitude to have regarding academia and am ever grateful for both.

%Academic help
Many other people also brought invaluable insight and help to this thesis:
%Andrew Jackson for the help
especially Andrew Jackson that was always around chatting about the latest Mortal Kombat test or a new fancy statistical move (or the other way around);
%Gavin Thomas especially for the help on TEM paper
Gavin Thomas for kindly spending time commenting my projects (especially Chapter \ref{chap:TEM_paper});
%Graeme Lloyd for the help on STD paper
Graeme Lloyd for the fast and enthusiastic responses to my questions (especially Chapter \ref{chap:STD_paper});
%Rich FitzJohn for coding skillz
Rich FitzJohn for enforcing his strict, yet invaluable, coding best practices (and for providing this thesis template);
%Trevor Hodkinson and Paula Murphy for steering committee
Trevor Hodkinson and Paula Murphy for following me through the thesis Committee;
%April Wright, Nick Matzke and Dave Bapst and the people from the Macro Journal Club
Dave Bapst, Nick Matzke, April Wright and the people from the Macro JC for providing a fertile ground for new ideas.

%Academic support
A great thank also to my office mates, Sive Finlay (PhD deirfi\'{u}r), Kevin Healy, Adam Kane and ``Marvin'' Qiang Yang for kindly supporting my random Frenchings.
Excellent technical and social support also came from the School of Natural Sciences staff, post-docs and students that made me feel at home since day one in Ireland:
%All alphabetically
Mirjam Ansorge, Yvonne Buckley, Joe Colgan, Luca Coscieme, Shaun Coutts, Ian Donohue, Alain Finn, Dave Kelly, Se\'{a}n Kelly, Deirdre McClean, Keith MacMahon, Conor Owens, Maureen Williams, Mike Whitfield and all the NERD club members.

%French people
Also, I'd like to thank people from the Universit\'{e} Montpellier 2 that are to be accounted for much in my motivation to undertake a PhD in the first place:
Julien Claude, Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery and Laurent Marivaux;
and, of course, Franck Barbi\`{e}re, S\'{e}bastien Enault, Morgan Le Ny, Sylvain Monteux, Benjamin Ramassamy and all the other friends from the two years masters as well.

%Family
And finally, I'd like to thank my folks, Marja, Jean-Philippe, Nina and Daan for not always understanding what all that was about (and asking me to explain more clearly) but still always being supporting my decision to undertake long and non-applied studies.

%Funding - TG: Kind of makes it not fit in the page... Don't really like it and Adam + Kev didn't acknowledged them and had no comments about not doing it.
This PhD was funded by the Trinity Postgraduate Research Studentship and the faculty of science engineering and maths and the European Commission CORDIS Seventh Framework Programme 772 (FP7) Marie Curie CIG grant (proposal number: 321696) provided funding for the research expenses.


