\chapter*{Specific chapter contributions and acknowledgements}
\addcontentsline{toc}{chapter}{Additional informations for each specific chapters}

\section*{Contributions}
For each chapters, I designed the experiments, ran the analysis, interpreted the results and wrote the manuscripts. Natalie Cooper helped for designing the experiment and writing the manuscript.

\section*{Data availability and reproducibility}
\subsection*{Effects of missing data on topological inference using a Total Evidence approach}
All the code used in this analysis is available on \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{GitHub} (\href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{    
goo.gl/4djNUf}) with some information on how to use the various functions. Additionally all the simulated data is available on \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{FigShare} (dx.doi.org/10.6084/m9.figshare.1306861).

\subsection*{Missing data in living mammals}
All data and analysis code is available on GitHub (\url{https://github.com/TGuillerme/Missing_living_mammals}).

\subsection*{Cretaceous-Palaeogene extinction does not affect mammalian disparity}
Data will be available on Dryad or Figshare.
Code for reproducing the analysis is available on GitHub (\url{https://github.com/TGuillerme/SpatioTemporal_Disparity}).


\section*{Specific chapters acknowledgements}
\subsection*{Effects of missing data on topological inference using a Total Evidence approach}
Thanks to Gavin Thomas, Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson, Andrew Jackson, Nick Matzke, and April Wright for useful comments on our simulation protocol and manuscript. Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster. Thanks to the two anonymous reviewers for their useful and enthusiastic comments.

\subsection*{Missing data in living mammals}
Thanks to David Bapst, Graeme Lloyd, Nick Matzke and April Wright.

\subsection*{Cretaceous-Palaeogene extinction does not affect mammalian disparity}
Thanks to Graeme Lloyd, Andrew Jackson, Gavin Thomas and Sive Finlay.



% \section*{Effects of missing data on topological inference using a Total Evidence approach}
% \subsection*{Data availability and reproducibility}
% All the code used in this analysis is available on \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{GitHub} (\href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{    
% goo.gl/4djNUf}) with some information on how to use the various functions. Additionally all the simulated data is available on \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{FigShare} (dx.doi.org/10.6084/m9.figshare.1306861).
% \subsection*{Acknowledgements}
% Thanks to Gavin Thomas, Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson, Andrew Jackson, Nick Matzke, and April Wright for useful comments on our simulation protocol and manuscript. Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster. Thanks to the two anonymous reviewers for their useful and enthusiastic comments.

% \section*{Missing data in living mammals}
% \subsection*{Data availability and reproducibility}
% All data and analysis code is available on GitHub (\url{https://github.com/TGuillerme/Missing_living_mammals}).
% \subsection*{Acknowledgements}
% Thanks to David Bapst, Graeme Lloyd, Nick Matzke and April Wright.

% \section*{Cretaceous-Palaeogene extinction does not affect mammalian disparity}
% \subsection*{Data availability and reproducibility}
% Data will be available on Dryad or Figshare.
% Code for reproducing the analysis is available on GitHub (\url{https://github.com/TGuillerme/SpatioTemporal_Disparity}).
% \subsection*{Acknowledgements}
% Thanks to Graeme Lloyd, Andrew Jackson, Gavin Thomas and Sive Finlay.